---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
  [![Travis build status](https://travis-ci.com/trottieralex/DHScalendR.svg?branch=master)](https://travis-ci.org/trottieralex/DHScalendR)
  <!-- badges: end -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# DHScalendR

The goal of DHScalendR is to make the use of the DHS calendar easier to use in R. **This package is currently being developped**, but it already contains functions to:

* Deal with CMC date format (back and forth)
* Split Calendar into months columns
* Recoding of the calendar
* Subset using dates (CMC)
* Check for combination of events (e.g. postpartum FP use = Birth & any contraception)

Next additions will include ways to:

* Locate events
* *NB. Also need better documentation and further tests.*
* *NB. The type of input will also evolve by version 0.3.0*


## Installation

<!-- You can install the released version of DHScalendR from [CRAN](https://CRAN.R-project.org) with: -->

To install the current version of this package:

```{r, eval=FALSE}
devtools::install_github("trottieralex/DHScalendR")
```


## Data

This package contains a small exemple dataset extracted from a DHS model dataset, and also a codebook to recode the calendar. The later is primarily to use with the calendar_recode function, but is available for consultation or use.

```{r, eval=FALSE}
# Exemple dataset
data <- dhs_extract
head(data)
?dhs_extract

# Calendar codebook
head(dhs_calendar_codebook)
```


## Transform CMC date format

The DHS widely use CMC format date (number of month counting from January 1900). A format often convenient for analysis, but hardly interpretable. These function make the transfer CMC from/to Date and years as decimal straigthforward. 

```{r echo=TRUE}
library(DHScalendR)

# The Date of the Teharan declaration:
cmc_to_Date(761)
cmc_from_Date(as.Date("1963-05-13")) #Show warning the 1st time

# or for April 19, 1989
cmc_to_Year(1072)
cmc_from_Year(1989.296)

```


## Split and recode DHS calendar

The split function divides the DHS calendar from a data frame into month columns that are added to the said data frame under the name: "CMCx" + [CMC date of the month].
Once split, the recode function coud be use to obtain more meaning full labels or classification.
```{r}
dataset <- dhs_extract[95:100, c("V008", "V025", "VCAL.1")]

#Dividing months into columns
dataset <- calendar_split(dataset, "V008", "VCAL.1")
(dataset <- dataset[ ,1:10])

#Recoding of calendar months
calendar_recode(dataset$CMCx1326)
calendar_recode(c("1","0","P","P","P","T","A","A","A","A","3","3","3"), 
                recoding = "modern_DHS6")

```

## Subset by dates

A simple way to subset the DHS calendar for a predefined period of time.

```{r}

dataset <- DHScalendR::dhs_extract[5:9, ]

# Request first half of 2014
calendar_subset(calendar = dataset$VCAL.1, 
                interview = dataset$V008, 
                starting = 1381, 
                ending = cmc_from_Date(as.Date("2015-06-30")))

```


## Check for combination of events

Look for any event2 that hapenned X months or less after any event1.

```{r}

example <- c("333000TPPP", "000000BPPP", "0000LLLBPPP")
fp <- dhs_calendar_codebook$Recode[dhs_calendar_codebook$modern_DHS6 == "Modern"]

# Check if use any modern FP in the 6 months after end pregnancy
calendar_combo(calendar = example, 
               event1 = c("B", "T"), 
               n_months = 6,
               event2 = fp) 
```

