---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# DHScalendR

<!-- badges: start -->
<!-- badges: end -->

The goal of DHScalendR is to make the use of the DHS calendar easier to use in R. *This package is currently being developped*, but it already contains functions to:

* Deal with CMC date format (back and forth)

Next additions will include ways to:

* Decomposition of the calendar string to months
* Recoding of the calendar
* Measure of specified events (i.e. contraceptive failure or post-partum adoption)


## Installation

<!-- You can install the released version of DHScalendR from [CRAN](https://CRAN.R-project.org) with: -->

```{r example}
# library(DHScalendR)
devtools::install_github("<<<<ADD info here")
```

## Transform CMC date format

The DHS widely use CMC format date (number of month counting from January 1900). A format often convenient for analysis, but hardly interpretable. These function make the transfer CMC from/to Date and years as decimal straigthforward. 

```{r example}

# The Date of the Teharan declaration:
cmc_to_Date(761) #1963-05-01
cmc_from_Date(as.Date("1963-05-13")) #761

# or for April 19, 1989
cmc_to_Year(1072) #1989.25
cmc_from_Year(1989.296) #1072

```

